---
title: "Data Analysis and Lab_Lab 3"
author: "Na SeungChan"
date: "`r Sys.Date()`"
mainfont : NanumGothic
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mdsr)
```




# ggplot2




## Scales

we will use `diamonds` dataset in ggplot2

```{r}
ggplot(diamonds, aes(carat, price)) +
    geom_bin2d()
#geom_bin2d counts its numbers in some block
ggplot(diamonds, aes(log10(carat), log10(price))) +geom_bin2d()
ggplot(diamonds, aes(carat, price)) +
    geom_bin2d() +
    scale_x_log10() +
    scale_y_log10()
ggplot(diamonds, aes(carat, price)) +
  geom_bin2d() + 
  scale_x_continuous(trans = "log10") + 
  scale_y_continuous(trans = "log10")
```


Of course, the below three graphs are all same.


### `RColorBrewer` Packages

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv))

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv)) +
  scale_colour_brewer(palette = "Set1")
```




# Statistical transform


bar chart? 근본적으로 다른 그래프. geom_bar는 우선 각 data를 그룹별로 구분된 작은 dataframe으로 만들고(by stat::count()), 두 번째 과정에서는 이를 통해 그래프를 그림.


만약 '비율'을 보고 싶다면? y에 비율(..prob..)을 써 주면 됨. 

geom_bar function을 쓰지 않는다면 그래프 그리는 과정에서 알파벳순으로 그래프가 그려지는 등 문제 발생.



# A grammer for data graphics


```{r}

```


`cut()` function makes the continuous variables into discrete.


```{r}
library(NHANES)
library(forcats)
```





if you want to make time-series graphs, it is recommended to use line-connected graphs(geom_line)



# Babynames


Notice that the name_plot object does not contain any contextual information, updated in the last code chunk. This was intentional, since we can update the data argument of name_plot and obtain an analogous plot for another name. This functionality makes use of the special %+% operator.




