---
title: "R_Basics_packages"
author: "Na SeungChan"
date: "`r Sys.Date()`"
mainfont : NanumGothic
monofont : UnShinmun
header-includes:
  - \usepackage{fvextra}
  - \fvset{breaklines}
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = 'center',
  fig.width = 7,
  fig.height = 5,
  out.width = '80%',
  dpi = 300
)
```

---

# Basic Packages

## tidyverse

```{r}
#install.packages('tidyverse')
```


```{r}
library(tidyverse)
comment = '
The tidyverse is a set of packages that work in harmony because they share common data representations and API design. 
The tidyverse package is designed to make it easy to install and load core packages from the tidyverse in a single command.
https://tidyverse.tidyverse.org/index.html
'
```


`library(tidyverse)` will load the core tidyverse packages:
* `ggplot2`, for data visualisation.
* `dplyr`, for data manipulation.
* `tidyr`, for data tidying.
* `readr`, for data import.
* `purrr`, for functional programming.
* `tibble`, for tibbles, a modern re-imagining of data frames.
* `stringr`, for strings.
* `forcats`, for factors.
* `lubridate`, for date/times.



## glue

```{r}
#install.packages('glue')
```


```{r}
library(glue)
comment = '
The glue package contains functions for string interpolation: 
gluing together character strings and R code.
https://glue.tidyverse.org/index.html
'
```






## MASS

```{r}
#install.packages('MASS')
```

```{r}
library(MASS)
comment = '

'
```

The `MASS` package has `select()` function, which conflicts with `dplyr::select()`. Generally, it is recommended to load `tidyverse` after loading `MASS`, because the `select()` function in `MASS` package is not used not so often.





























