---
title: "Practice_R_20220927"
author: "Na SeungChan"
date: "2022-09-27"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mdsr)
```

## R Basics

```{r as a calculator}
3+2 #addition
3-2 #subtraction
3*2 #multiplication
3/2 #division
3^2 #power cf) python : 3**2
5%%2 #Remainder
5%/%2 #quotient
0/0 #illegal arithmetic yields a NaN(Not a Number)
class(NaN)
```


```{r variables, functions}
aaa <- c(1,2) #preferred.
bbb = c(1,2) #avoid.
c(1,2) -> ccc #weird.

#form of function_name(arg1 = val1, arg2 = val2) : 

x <- seq(1,10,length.out = 5)


```

## Working Directory

```{r}
getwd()

aaa <- read.csv("candidate-votes-by-voting-place-4.csv", header = F)
aaa
```


## R markdown Basics

### texts

\*italic\* or \_italic\_
*italic*
_italic_

\*\*bold\*\* or \_\_bold\_\_
**bold**
__bold__

\`code\`
`code`

superscript \^super\^
super^2^

subscript \~sub\~
sub~2~

\#1st level header
\#\#2nd level header

* Bulleted list item 1
* item 2
  *item 2a(by \\t)

1. Numbered list item 1
0. item 2. the number is not important.

link : 
<domain>
[text](domain)
![imgname](./imgfilename)

Tables

| 1st Header | 2nd Header | 3rd Header | 4st Header |
| : -------- | ---------- | ----------:|:--------- : |
| con        | con | con | con |
| con        | con | con | con |


inline code chunk : `r 1+1` will embed the result of R code directly into the test. 

for example, `r 4+5` is that.

ss



### code chunk

eval = FALSE : prevents code from being evaluated
include = FALSE : runs the code, but don't show the code or results
echo = FALSE : only prevents the code chunk from appearing
error = True : cause the render to continue even if codes returns an error.
cache = TRUE : will save the output of the chunk on disc. note that even if the '.csv' files in code change, the code chunk does not rerun.
message = FALSE/warning = FALSE : do not show the messages or warnings.
results = "hide" or fig.show = "hide"

knitr::opts_chunk$set(echo = TRUE) - global options




```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
