---
title: "R_basic_statistics"
author: "Na SeungChan"
date: "`r Sys.Date()`"
mainfont : NanumGothic
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




# Week2 : 

```{r}
gender <- factor(c(rep('M', 8), rep('F', 10)))
score <- c(98, 90, 96, 54, 43, 87, 88, 90, 94, 92, 81, 79, 85, 91, 79, 88, 89, 83)

```


## 일변량 자료의 요약

### 그래프프

```{r : histogram}
hist(score)
hist(score, freq = F)
hist(score, breaks = 20)
```


```{r : stem}
stem(score)
stem(score, scale = 0.1)
```



```{r : boxplot}
boxplot(score)
boxplot(score ~ gender)
boxplot(score ~ gender, horizontal = TRUE)
``````


### 수치통계량

```{r}

```




## 다변량 자료의 요약

```{r}
math <- c(66, 64, 48, 46, 78, 60, 90, 50, 66, 70)
phy <- c(70, 68, 46, 48, 84, 64, 92, 52, 68, 72)
```


### 그래프

```{r}
plot(math, phy, main = 'tile', sub = 'nyan', xlim = c(45, 85), ylab = 'se', type = 'b')
```




# week 3

## vector

```{r}
x <- c(1, 2, 3, 5, 6)
x[1]
x <- 5:10
x
x <- rep(c(1, 2, 3), each = 5)
x
x <- rep(c(1, 2, 4), 5)
x
x <- seq(1, 10, 2)
x
x <- seq(1, 10, length = 5)
x
rm(x)
x
x <- sample(c(1, 2, 4), size = 5, replace = TRUE)
x
```

## Matrix

```{r}
x1 <- c(1, 2, 3, 4, 5)
x2 <- c(6, 7, 8, 9, 10)
x3 <- seq(11, 15)
x4 <- sample(c(1, 2, 4), size = 5, replace = TRUE)

X <- rbind(x1, x2, x3, x4)
X
X <- cbind(x1, x2, x3, x4)
X
```


# Week4

## 정규분포


```{r}
x <- seq(-3, 3, by = 0.01)
fx <- dnorm(x, mean = 0, sd = 1) #확률밀도함수
plot(x, fx, type = 'l', main = 'pdf of normal distribution')
px <- pnorm(x, mean = 0, sd = 1, lower.tail = TRUE)
plot(x, px, type = 'l', main = 'cdf of normal distribution')
qx <- qnorm(px, mean = 0, sd = 1, lower.tail = TRUE)
```



```{r}
par(mfrow = c(1, 3))
plot(x, fx, type = 'b', main = 'pdf of normal distribution')
plot(x, px, type = 'l', main = 'cdf of normal distribution')
plot(px, qx, type = 'l', main = 'quantile for px')
dev.off()
```


## 이항분포


```{r}
x <- 0:10
fx <- dbinom(x, size = 10, prob = 0.3)
plot(x, fx, type = 'h', main = 'pdf of normal distribution') #막대형 히스토그램
px <- pbinom(x, size = 10, prob = 0.3)
plot(x, px, type = 'h', main = 'cdf of normal distribution')
```


## 카이제곱분포


```{r}
x <- seq(0, 10, by = 0.01)
fx <- dchisq(x, df = 12)
fx
```


## 난수 생성

```{r}
s1 <- sample(c(1, 2, 4), size = 5, replace = TRUE)
s2 <- runif(n = 10, min = 0, max = 2) #random from uniform
```



## qqplot

```{r}
qqnorm()
qqline()
```
















































































































































