---
title: "Data Analysis and Lab_Lab 3"
author: "Na SeungChan"
date: "`r Sys.Date()`"
mainfont : NanumGothic
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mosaic)
library(mdsr)
library(MASS)
```



# Exercise 1


## (a) Create a scatter plot of each person’s height against their father’s height


```{r}
ggplot(data = Galton) +
  geom_point(mapping = aes(x = father, y = height))
```



## (b) Separate your plot into facets by sex


```{r}
ggplot(data = Galton) +
  geom_point(mapping = aes(x = father, y = height)) +
  facet_wrap(~sex)
```



## (c) Add regression lines to all of your facets; use `geom_smooth(method = "lm")`.


```{r}
ggplot(data = Galton) +
  geom_point(mapping = aes(x = father, y = height)) +
  facet_wrap(~sex) +
  geom_smooth(mapping = aes(x = father, y = height), method = "lm", formula = y ~ x)
```



## (d) Analyze the results of (a), (b) and (c) in 2 ~ 3 sentences.


산점도상 자녀의 키와 아버지의 키는 그래프로 봤을 때 양의 상관관계를 가지는 것으로 보인다. 이와 같은 결과는 '남성과 여성의 키 차이를 고려'하여 자녀의 성별에 따라 서로 다른 facet에 각 그래프를 그려도 동일하게 나타난다. 실제 이에 따라 회귀직선을 mapping해 본 경우에도 비슷한 결과가 나온다.



# Exercise 2


```{r eval = FALSE}
MLB_teams
```



## (a) See how many variables you can illustrate on a single plot in R. The current record is 7. [Note: this is not good graphical practice—it is merely an exercise to help you understand how to use visual cues and aesthetics!]


```{r}
MLB_teams %>% 
  mutate(z_pay = (payroll-mean(payroll))/sd(payroll)) %>%
  ggplot() +
  geom_point(mapping = aes(x = z_pay, y = WPct, color = factor(yearID))) +
  xlim(-4,4) +
  xlab('z-score for their payroll') +
  ylab('Win Percent') +
  labs(title = 'WPct vs. Payroll')
```



## (b) Use the `MLB_teams` data in the `mdsr` package to create an informative data graphic that illustrates the relationship between winning percentage `WPct` and payroll `payroll` in context.


```{r}
ggplot(data = MLB_teams, mapping = aes(x = payroll, y = WPct)) +
  geom_point() +
  geom_smooth(method = "lm")
```


Overall there is a positive correlation between payroll and WPct. However, overall payroll tends to increase as time goes.

```{r}
MLB_teams %>% 
  group_by(yearID) %>% 
  summarise(sum(payroll)) 
```


So, we will use z-score. In z-scores, we can see teams below -2.0 did not exist, but teams with z-scores over 3.0 was 3 teams. 

```{r}
MLB_teams %>% 
  mutate(z_pay = (payroll-mean(payroll))/sd(payroll)) %>%
  ggplot() +
  geom_point(mapping = aes(x = z_pay, y = WPct, color = factor(yearID))) +
  xlim(-4,4) +
  xlab('z-score for their payroll') +
  ylab('Win Percent') +
  labs(title = 'WPct vs. Payroll')
```


 
# Exercises 3 : Write a code to create a data object named `Binary_medv` whose value is `"rich"` if the value of `medv` is greater than 25, `"not so"` if not. Use both `with()` and `ifelse()`.


```{r}
my_Boston <- Boston %>%
  mutate(Binary_medv = with(Boston, ifelse(medv > 25, 'rich', 'not so'))) %>%
  relocate(Binary_medv)
head(my_Boston)
```


