---
title: "Data Analysis and Lab."  
subtitle: 'Lab 11: Splines & GAM'
author: "Suhwan Bong"
date: "2023-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

In this lab, we re-analyze the `Wage` data considered in the examples through- out this chapter, in order to illustrate the fact that many of the complex non-linear fitting procedures discussed can be easily implemented in R. We begin by loading the ISLR library, which contains the data.
```{r}
library(ISLR)
attach(Wage)
head(Wage)
```

## 1. Splines


In order to fit regression splines in R, we use the splines library. We saw that regression splines can be fit by constructing an appropriate matrix of basis functions. The `bs()` function generates the entire matrix of basis functions for splines with the specified set of knots. By default, cubic splines are produced. Fitting `wage` to `age` using a regression spline is simple:

```{r}
library(splines)
fit <- lm(wage ~ bs(age, knots = c(25,40,60)), data = Wage)
agelims <- range(age)
age.grid <- seq(from=agelims[1],to=agelims[2])
pred <- predict(fit, newdata = list(age=age.grid), se = T)
plot(age, wage, col = "gray")
lines(age.grid, pred$fit, lwd = 2)
lines(age.grid, pred$fit + 2 * pred$se, lty = "dashed")
lines(age.grid, pred$fit - 2 * pred$se, lty = "dashed")
```

Here we have prespecified knots at ages 25, 40, and 60. This produces a spline with six basis functions. (Recall that a cubic spline with three knots has seven degrees of freedom; these degrees of freedom are used up by an intercept, plus six basis functions.) We could also use the `df` option to produce a spline with knots at uniform quantiles of the data.

```{r}
dim(bs(age, knots = c(25, 40, 60)))
dim(bs(age, df = 6))
attr(bs(age, df = 6), "knots")
```

In this case `R` chooses knots at ages 33.8, 42.0, and 51.0, which correspond to the 25th, 50th, and 75th percentiles of `age`. The function `bs()` also has a `degree` argument, so we can fit splines of any degree, rather than the default degree of 3 (which yields a cubic spline).

In order to instead fit a natural spline, we use the `ns()` function. Here we fit a natural spline with four degrees of freedom.
```{r}
fit2 <- lm(wage ~ ns(age,df=4),data=Wage)
pred2 <- predict(fit2,newdata=list(age=age.grid),se=T)
plot(age, wage, col = "gray")
lines(age.grid, pred2$fit,col="red",lwd=2)
```


As with the `bs()` function, we could instead specify the knots directly using the `knots` option. In order to fit a smoothing spline, we use the `smooth.spline()` function.

```{r}
plot(age,wage,xlim=agelims ,cex=.5,col="darkgrey")
title (" Smoothing Spline ")
fit=smooth.spline(age,wage,df=16)
fit2=smooth.spline(age,wage,cv=TRUE)
fit2$df
lines(fit,col="red",lwd=2)
lines(fit2,col="blue",lwd=2)
legend("topright",legend=c("16 DF","6.8 DF"), col=c("red","blue"),lty=1,lwd=2,cex=.8)
```
Notice that in the first call to `smooth.spline()`, we specified `df=16`. The function then determines which value of $\lambda$ leads to 16 degrees of freedom. In the second call to `smooth.spline()`, we select the smoothness level by cross-validation; this results in a value of $\lambda$ that yields 6.8 degrees of freedom. In order to perform local regression, we use the `loess()` function.

```{r}
plot(age,wage,xlim=agelims ,cex=.5,col="darkgrey")
title (" Local Regression ")
fit=loess(wage~age,span=.2,data=Wage)
fit2=loess(wage~age,span=.5,data=Wage)
lines(age.grid,predict(fit,data.frame(age=age.grid)), col="red",lwd=2)
lines(age.grid,predict(fit2,data.frame(age=age.grid)), col="blue",lwd=2)
legend("topright",legend=c("Span=0.2","Span=0.5"), col=c("red","blue"),lty=1,lwd=2,cex=.8)
```


Here we have performed local linear regression using spans of 0.2 and 0.5: that is, each neighborhood consists of 20 % or 50 % of the observations. The larger the span, the smoother the fit. The `locfit` library can also be used for fitting local regression models in `R`.

## 2. GAM
We now fit a GAM to predict `wage` using natural spline functions of `year` and `age`, treating `education` as a qualitative predictor. Since this is just a big linear regression model using an appropriate choice of basis functions, we can simply do this using the `lm()` function.
```{r}
gam1=lm(wage~ns(year ,4)+ns(age ,5)+education ,data=Wage)
```

We now fit the model using smoothing splines rather than natural splines. In order to fit more general sorts of GAMs, using smoothing splines or other components that cannot be expressed in terms of basis functions and then fit using least squares regression, we will need to use the `gam` library in `R`.

The `s()` function, which is part of the `gam` library, is used to indicate that we would like to use a smoothing spline. We specify that the function of `year` should have 4 degrees of freedom, and that the function of `age` will have 5 degrees of freedom. Since `education` is qualitative, we leave it as is, and it is converted into four dummy variables. We use the `gam()` function in order to fit a GAM using these components. All of the terms are fit simultaneously, taking each other into account to explain the response.

```{r}
library(gam)
gam.m3=gam(wage~s(year ,4)+s(age ,5)+education ,data=Wage)
par(mfrow=c(1,3))
plot(gam.m3, se=TRUE,col="blue")
```

The generic `plot()` function recognizes that `gam.m3` is an object of class `gam`, and invokes the appropriate `plot.Gam()` method. Conveniently, even though `gam1` is not of class `gam` but rather of class `lm`, we can still use `plot.Gam()` on it.

```{r}
plot.Gam(gam1, se=TRUE, col="red")
```

Notice here we had to use `plot.Gam()` rather than the generic `plot()` function. In these plots, the function of `year` looks rather linear. We can perform a series of ANOVA tests in order to determine which of these three models is best: a GAM that excludes `year` (M1), a GAM that uses a linear function of `year` (M2), or a GAM that uses a spline function of `year` (M3).

```{r}
gam.m1=gam(wage ~ s(age ,5)+education ,data=Wage)
gam.m2=gam(wage ~ year+s(age ,5)+education ,data=Wage)
anova(gam.m1,gam.m2,gam.m3,test="F")
```

We find that there is compelling evidence that a GAM with a linear function of `year` is better than a GAM that does not include `year` at all (p-value = 0.00014). However, there is no evidence that a non-linear function of `year` is needed (p-value = 0.349). In other words, based on the results of this ANOVA, M2 is preferred. The `summary()` function produces a summary of the gam fit.

```{r}
summary(gam.m3)
```

The p-values for `year` and `age` correspond to a null hypothesis of a linear relationship versus the alternative of a non-linear relationship. The large p-value for `year` reinforces our conclusion from the ANOVA test that a linear function is adequate for this term. However, there is very clear evidence that a non-linear term is required for `age.` We can make predictions from `gam` objects, just like from `lm` objects, using the `predict()` method for the class `gam`. Here we make predictions on the training set.

```{r}
preds = predict(gam.m2, newdata = Wage)
```

We can also use local regression fits as building blocks in a GAM, using the `lo()` function.

```{r}
gam.lo=gam(wage ~ s(year,df=4)+lo(age,span=0.7)+education, data=Wage)
plot.Gam(gam.lo, se=TRUE, col="green")
```

Here we have used local regression for the age term, with a span of 0.7. We can also use the `lo()` function to create interactions before calling the `gam()` function. For example,

```{r}
gam.lo.i=gam(wage ~ lo(year,age,span=0.5)+education, data=Wage)
```

fits a two-term model, in which the first term is an interaction between `year` and `age`, fit by a local regression surface. We can plot the resulting two-dimensional surface if we first install the `interp` package.
```{r}
library(interp)
plot(gam.lo.i)
```

In order to fit a logistic regression GAM, we once again use the `I()` function in constructing the binary response variable, and set `family=binomial`.

```{r}
gam.lr=gam(I(wage>250) ~ year+s(age,df=5)+education, family=binomial,data=Wage)
par(mfrow=c(1,3))
plot(gam.lr,se=T,col="green")
```

It is easy to see that there are no high earners in the `<HS` category:
```{r}
table(education ,I(wage >250))
```
Hence, we fit a logistic regression GAM using all but this category. This provides more sensible results.

```{r}
gam.lr.s=gam(I(wage>250) ~ year+s(age,df=5)+education,family= binomial,data=Wage,subset=(education!="1. < HS Grad"))
plot(gam.lr.s,se=T,col="green")

```


## 3. Exercises

### Problem 1. 
This question uses the variables `dis` (the weighted mean of distances to five Boston employment centers) and `nox` (nitrogen oxides concentration in parts per 10 million) from the `Boston` data. We will treat `dis` as the predictor and `nox` as the response.

```{r}
library(MASS)
head(Boston)
```


(a) Use the `poly()` function to fit a cubic polynomial regression to predict `nox` using `dis`. Report the regression output, and plot the resulting data and polynomial fits.
(b) Plot the polynomial fits for a range of different polynomial degrees (say, from 1 to 10), and report the associated residual sum of squares.
(c) Perform cross-validation or another approach to select the optimal degree for the polynomial, and explain your results.
(d) Use the `bs()` function to fit a regression spline to predict `nox` using `dis`. Report the output for the fit using four degrees of freedom. How did you choose the knots? Plot the resulting fit.
(e) Now fit a regression spline for a range of degrees of freedom, and plot the resulting fits and report the resulting RSS. Describe the results obtained.
(f) Perform cross-validation or another approach in order to select the best degrees of freedom for a regression spline on this data. Describe your results.

### Problem 2.
This question relates to the `College` data set.

```{r}
head(College)
```


(a) Split the data into a training set and a test set. Using out-of-state tuition as the response and the other variables as the predictors, perform forward stepwise selection on the training set in order to identify a satisfactory model that uses just a subset of the predictors.
(b) Fit a GAM on the training data, using out-of-state tuition as the response and the features selected in the previous step as the predictors. Plot the results, and explain your findings.
(c) Evaluate the model obtained on the test set, and explain the results obtained.
(d) For which variables, if any, is there evidence of a non-linear relationship with the response?