---
title: "Data Analysis and Lab HW4"
author: "Na SeungChan"
date: "`r Sys.Date()`"
mainfont : NanumGothic
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MatchIt)
library(WeightIt)
library(sandwich)
library(optmatch)
library(cobalt)
library(marginaleffects)
```



# Problem 1.

```{r}
df <- read.csv("welder.csv")
head(df)
```



---

## (a)

Compare the covariate balance between treated group and control group. Find the mean of treated group, mean of control group, and standardized mean difference for each covariate. Discuss the result.


```{r}
no.adj <- matchit(welder ~ age + race + smoker, data = df, method = NULL, distance = "glm")
summary(no.adj)

df %>%
  group_by(welder) %>%
  summarise(mean(DPC), sd(DPC))
```

Standardized Mean Difference for each covariate : 크게 차이나므로 이를 적절히 보정해야 한다.




---

## (b) 

Estimate the propensity score via logistic regression. Plot the density of propensity score for treated and control groups. Discuss the result.


```{r}
ps.glm <- glm(welder ~ age + race + smoker, data = df, family=binomial, x=T)
summary(ps.glm)
estimates_ps <- ps.glm$fitted.values
# 1/(1+exp((-1)*(model.matrix(ps.glm)[,-1]%*%ps.glm$coefficients[c(2:5)])))

hist(estimates_ps[df$welder==0], col = rgb(1, 0, 0, 0.2), xlab = "Propensity Score", main = "Treated (green) vs. Control (red)")
hist(estimates_ps[df$welder==1], col = rgb(0, 1, 0, 0.2), add = T)
```

propensity score는 treatment group의 경우와 control group의 경우, 즉 welder = 0과 1인 경우 각각 1과 0에 몰려 분포가 크게 차이난다.



---

## (c) 

Implement 1:1 Propensity Score Matching with replacement (use the propensity score estimated by logistic regression). Discuss the covariate balance (standardized mean difference) after matching.


```{r}
ps.mat11_7 <- matchit(welder ~ age + race + smoker, data = df, method = 'nearest', distance = estimates_ps, replace = T)
summary(ps.mat11_7, un = F)
ps.mat11_7$match.matrix
```


matching 후, 모든 Std. Mean Diff의 최댓값이 0.1보다 약간 큰 수준으로acceptable하다.  이와 같이 group 간 차이가 보정되어 각 group의 차이가 없는 것으로 생각하고 각 group의 DPC value 차이를 treatment가 random assignment된 것에 따라 발생한 인과관계에서 발생하는 ATT로 생각할 수 있게 된다.

참고로, 아래의 결과를 확인하면 `replace = T` option이 얼마나 큰 차이를 만들었는지 알 수 있다. 41이 얼마나 많이 반복되는지! sample 수가 아무리 작아도 이렇게 하는 게 현실 연구에서도 맞는지는 모르겠다...



---

## (d)

Using the matched set in (c), find the estimate and 95% confidence interval of the average treatment effect on treated (ATT).


```{r}
mat117_data <- match.data(ps.mat11_7)
fit_DPC_treat <- lm(DPC ~ welder, data = mat117_data, weights = weights)
avg_comparisons(fit_DPC_treat, variables = "welder", vcov = 'HC3', newdata = subset(mat117_data, welder == 1), wts = "weights")
```

95% CI : (-0.371, 1.35)
ATT estimate : 0.489



---

## (e) 

Construct a propensity score stratification with 3 strata (use the propensity score estimated by logistic regression). Discuss the covariate balance across subclasses.


```{r}
ps.mat11_3 <- matchit(welder ~ age + race + smoker, data = df, methods = 'nearest', distance = estimates_ps, subclass = 3)
summary(ps.mat11_3, un = F)

mat113_data <- match.data(ps.mat11_3)
fit_DPC_treat2 <- lm(DPC ~ welder, data = mat113_data, weights = weights)
avg_comparisons(fit_DPC_treat2, variables = "welder", vcov = 'HC3', newdata = subset(mat113_data, welder == 1), wts = "weights")
```

replace = FALSE로 설정한 결과 standardized mean diff가 매우 크다. 이걸 쓰는 게 적절하지 않아 보인다.



---

## (f) 

Using the strata in (e), find the estimate and 95% confidence interval of the average treatment effect (ATE).


```{r}
ps.mat12_3 <- matchit(welder ~ age + race + smoker, data = df, methods = 'nearest', distance = 'glm', link = 'probit', subclass = 3, replace = T)
summary(ps.mat12_3, un = F)

mat123_data <- match.data(ps.mat12_3)
fit_DPC_treat3 <- lm(DPC ~ welder, data = mat123_data, weights = weights)
avg_comparisons(fit_DPC_treat3, variables = "welder", newdata = subset(mat123_data, welder == 1), wts = "weights")
```

(-0.298, 1.33), 0.515.


estimand option 'ATE'가 몇몇 option에 대해 가능하지만... 다음과 같은 오류를 출력한다.


```{r, error=TRUE}
ps.mat12_3 <- matchit(welder ~ age + race + smoker, data = df, methods = 'nearest', distance = 'glm', link = 'probit', estimand = 'ATE', subclass = 3, replace = T)
summary(ps.mat12_3, un = F)

mat123_data <- match.data(ps.mat12_3)
fit_DPC_treat3 <- lm(DPC ~ welder, data = mat123_data, weights = weights)
avg_comparisons(fit_DPC_treat3, variables = "welder", newdata = subset(mat123_data, welder == 1), wts = "weights")
```


---

## (g) 

Find the estimate and 95% confidence interval of the average treatment effect (ATE) using inverse probability weighting (IPW, use the propensity score estimated by logistic regression).


```{r}
Wmodel <- weightit(welder ~ age + race + smoker, data = df, methods = 'nearest', estimand = 'ATE', distance = estimates_ps)
bal.tab(Wmodel, un = T)

W_data <- df
W_data$ipw.weights <- Wmodel$weights
fit_Wtreat <- lm(DPC ~ welder, data = W_data, weights = ipw.weights)
avg_comparisons(fit_Wtreat, variables = "welder", vcov = 'HC3', newdata = subset(W_data, welder == 1), wts = 'ipw.weights')
```

(0.0374, 1.15), 0.592


---

## (h) 

Compare three estimates in (d), (f), (g). Choose the estimate that you consider the best and discuss the reason.


우선, n이 적으므로 replace를 하지 않은 (f)는 사용하지 않는다. (f)의 model이 std.error이 적어 더 좋아 보인다.