---
title: "R for Data Science Study note_04 Data Transformation"
author: "Na SeungChan"
date: "2023-07-19"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
monofont: UnShinmun
mainfont: UnDotum
---

```{r setup, include=FALSE}
library('tidyverse')
library('nycflights13')

```

# Data Transformation


## dplyr basics


* Pick observations by their values (filter()).
* Reorder the rows (arrange()).
* Pick variables by their names (select()).
* Create new variables with functions of existing variables (mutate()).
* Collapse many values down to a single summary (summarise()).

These can all be used in conjunction with group_by() which changes the scope of each function from operating on the entire dataset to operating on it group-by-group. These six functions provide the verbs for a language of data manipulation.


Together these properties make it easy to chain together multiple simple steps to achieve a complex result. Let’s dive in and see how these verbs work.

```{r}
fly <- nycflights13::flights
```




## Use `filter()` to filter rows

filter() allows you to subset observations based on their values.


```{r}
filter(flights, month == 1, day == 1)
```

you can use comparision operator(==, !=, >, <, >=, <=) and logical operator(|, &, xor(,)) with filter. + you can use `%in%`


### Exercises

```{r}
#Q1
fly %>% filter(arr_delay >= 120)
fly %>% filter(dest == 'IAH'|dest == 'HOU')
fly %>% filter(carrier %in% c('UA', 'DL', 'AA'))
```


```{r}
#Q2
fly %>% filter(between(arr_delay, 50, 100))#includes 50 and 100. 
```


```{r}
fly %>%
  filter(is.na(dep_time)) %>%
  count()
```


## Arrange rows with `arrange()`

```{r}
fly %>%
  arrange(year)
fly %>%
  arrange(year, month)#default : ascending
fly %>%
  arrange(desc(year), desc(month))#NAs are always in the end.
```


### Exercises

```{r}
#Q1
arrange(fly, desc(is.na(dep_time)))
```



```{r}
#Q2
fly %>%
  arrange(desc(dep_delay))
```



## Select Columns with `select()`


```{r}
select(fly, year, month, day)
select(fly, -(year:day))
select(fly, !(year:day))
select(fly, year:day)
select(fly, c(1,3))
select(fly, seq(1,5,2))
```


```{r}
select(fly, starts_with('arr'))
select(fly, ends_with('delay'))
select(fly, contains('dep'))
select(fly, matches('._time'))#see strings
```







