---
title: "R for Data Science Study note_EDA"
author: "Na SeungChan"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
monofont: UnShinmun
mainfont: UnDotum
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse')
```

# EDA

The goal of EDA : Develop an understanding of my data.

making 'quality' = exploring a quantity of data

Two key question :

* What type of variation occurs within my variables?
    *
* What type of covariation occurs between my variables?
    *


## Variation 

def : the tendency of the values of a variable to change from m to m


```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
ggplot(data = diamonds) + 
  geom_histogram(mapping = aes(x = carat), binwidth = 0.5)
ggplot(data = filter(diamonds, carat < 3)) +
  geom_histogram(mapping = aes(x = carat), binwidth = 0.5)
ggplot(data = filter(diamonds, carat < 3)) +
  geom_freqpoly(mapping = aes(x = carat, color = cut), binwidth = 0.1)
```

Now we can recognize the variation. Then what should you look for in your plots? And what type of follow-up questions should you ask?


### Typical Values

the story of 'bar(discrete)/hist(continuous)'

  * Which value are the most common? Why?
  * Which values are rere? Why? Does that match my expectation?
  * Can I see any unusual patterns? What might explain them?


### Unusual Values

the story of 'outliers'


```{r}
ggplot(diamonds) + 
  geom_histogram(mapping = aes(x = y), binwidth = 0.5)
ggplot(diamonds) + 
  geom_histogram(mapping = aes(x = y), binwidth = 0.5) +
  coord_cartesian(ylim = c(0, 50))
```


### Missing Values

If you’ve encountered unusual values in your dataset, and simply want to move on to the rest of your analysis, you have two options.


1. Drop the entire row with the strange values:

```{r}
diamonds2 <- diamonds %>% 
  filter(between(y, 3, 20))
```


2. Replace the unusual values with missing values.

```{r}
diamonds2 <- diamonds %>% 
  mutate(y = ifelse(y < 3 | y > 20, NA, y))
#diamonds3 <- diamonds %>%
#  mutate(w = case_when(y < 3|y > 20 ~ NA, TRUE ~ y)) why dose not work? 'NA' makes r think I want logical vector.
diamonds4 <- diamonds %>%
  mutate(w = case_when(y < 3|y > 20 ~ NA_real_, TRUE ~ y))
```



### Exercises

```{r}
#Q1
ggplot(diamonds) +
  geom_histogram(mapping = aes(x = x))
ggplot(diamonds) +
  geom_histogram(mapping = aes(x = y))
ggplot(diamonds) +
  geom_histogram(mapping = aes(x = z))
ggplot(data = filter(diamonds, x<9, y<10, z<10)) +
  geom_point(mapping = aes(x = x, y = y, color = z)) +
  coord_fixed()
```


```{r}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price), binwidth = 1)
```


```{r}
diamonds3 <- diamonds %>%
  filter(carat == 0.99|carat == 1)
ggplot(diamonds3) + 
  geom_bar(mapping = aes(x = carat))
```



```{r}
ggplot(diamonds) + 
  geom_point(mapping = aes(x = carat, y = price)) +
  coord_cartesian(xlim = c(0, 5), ylim = c(0, 20000))
```



## Covariation

### Categorical vs. Continuous

```{r}
ggplot(data = diamonds, mapping = aes(x = price)) + 
  geom_freqpoly(mapping = aes(colour = cut), binwidth = 500)
ggplot(data = diamonds, mapping = aes(x = price, y = ..density..)) + 
  geom_freqpoly(mapping = aes(colour = cut), binwidth = 500)
```


```{r}
ggplot(data = diamonds, mapping = aes(x = cut, y = price)) +
  geom_boxplot()
```

```{r}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = reorder(class, hwy, FUN = median), y = hwy))
#reordering
```


```{r}
#Q1
a <- nycflights13::flights
ggplot(a) +
  geom_boxplot(mapping = aes(x = dep_time, group = is.na(dep_time)))
```


### 2 categorical

YOU need to count!

```{r}
ggplot(data = diamonds) +
  geom_count(mapping = aes(x = cut, y = color))
diamonds %>% 
  count(color, cut) %>%  
  ggplot(mapping = aes(x = color, y = cut)) +
    geom_tile(mapping = aes(fill = n))
```


### 2 Continuous

```{r}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price))#scatterplot becomes less useful
ggplot(data = diamonds) + 
  geom_point(mapping = aes(x = carat, y = price), alpha = 1 / 100)
ggplot(data = filter(diamonds, carat < 3)) +
  geom_bin2d(mapping = aes(x = carat, y = price))
ggplot(data = filter(diamonds, carat < 3), mapping = aes(x = carat, y = price)) + 
  geom_boxplot(mapping = aes(group = cut_width(carat, 0.1)))
ggplot(data = filter(diamonds, carat < 3), mapping = aes(x = carat, y = price)) + 
  geom_boxplot(mapping = aes(group = cut_number(carat, 20)))
```




## Patterns and models

If you spot a pattern, ask yourself:
* Could this pattern be due to coincidence (i.e. random chance)?
* How can you describe the relationship implied by the pattern?
* How strong is the relationship implied by the pattern?
* What other variables might affect the relationship?
* Does the relationship change if you look at individual subgroups of the data?

If two variables covary, you can use the values of one variable to make better predictions about the values of the second. If the covariation is due to a causal relationship (a special case), then you can use the value of one variable to control the value of the second.

modelr is the strongest way to deal with patterns.

```{r}
library(modelr)

mod <- lm(log(price) ~ log(carat), data = diamonds)

diamonds2 <- diamonds %>% 
  add_residuals(mod) %>% 
  mutate(resid = exp(resid))

ggplot(data = diamonds2) + 
  geom_point(mapping = aes(x = carat, y = resid))
```




