---
title: "R_Basics_GLM"
author: "Na SeungChan"
date: "`r Sys.Date()`"
mainfont : NanumGothic
monofont : UnShinmun
header-includes:
  - \usepackage{fvextra}
  - \fvset{breaklines}
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = 'center',
  fig.width = 7,
  fig.height = 5,
  out.width = '80%',
  dpi = 300
)

library(tidyverse)
```

---

# R Base

```{r}

#주석 
comment = '
If this story was the one to draw a miracle,
when time comes someday I will be meeting you again
'
```





```{r}
#help("list")
#?list
```




## Basic Operators

### Arithmetic Operators

from () to ().

```{r}
# addition
3+4
# subtraction
6-5
# multiplication
5*2
# division
7/3
# integer division
7%/%3
# remainder
7%%3
# power 
4**3
4^3
```


### Relational Operators

```{r}
# > : LHS is greater than RHS
4 > 7
9 > 0.6
8 > 8
# < : RHS is greater than LHS
4 < 7
9 < 0.6
8 < 8
```

```{r}
# >= : LHS is greater than or equal to RHS
4 >= 7
9 >= 0.6
8 >= 8
# <= : RHS is greater than or equal to LHS
4 <= 7
9 <= 0.6
8 <= 8
```


```{r}
# == : LHS is equal to RHS
4 == 7
9 == 0.6
8 == 8
# != : LHS is not equal to RHS
4 != 7
9 != 0.6
8 != 8
```

Floating point error : `all.equal(a, b)` function can solve this problem.


### Logical Operators

`&` : and 
`|` : or
`!` : not

```{r}
print(3&0)
print(TRUE&FALSE)
print(-11|2)
print(TRUE|FALSE)
print(!1)
print(!TRUE)
```

## printing

```{r}
s1 <- 'asdf'
s2 <- 'tmf'
s3 <- 'lly'
s4 <- 'kyobo'
s5 <- 4+5
print(s1)
print(paste0(s1, s3))
print(paste(s2, s5, sep=''))
cat(s3, s2)
```

```{r, eval=FALSE}
for (i in 1:3) {
  glue
}
```


## Basic Variables, data types

| Data Types | Description |
|---------|-----|
| numeric | integer/double |
| character | string/character |
| logical | TRUE/FALSE |

```{r}
a1 <- 5
a2 <- 2.5
a3 <- 5L
a4 <- 'a'
a5 <- 'asdf'
a6 <- TRUE

mode(a1)
mode(a2)
mode(a3)
mode(a4)
mode(a5)
mode(a6)

typeof(a1)
typeof(a2)
typeof(a3)
typeof(a4)
typeof(a5)
typeof(a6)
```


```{r}
is.logical(a1)
as.logical(a1)
```



`mode()` function returns the type of object. `typeof()` returns 'double' or 'integer'.


## data structures

| Dimension | Single Data Type | Multiple Data Types |
|-----|-----|-----|
| 1 | vector | list |
| 2 | matrix | data.frame |
| 3 | array | tibble* |

* `tibble` from `tibble` package. tibble supports `layer` dimension.


### vector

```{r}
v1 <- c(1, 2, 3, 5, 7)
v2 <- 6:10
v3 <- seq(from=10, to=1, by=-4)
v4 <- c(rep(c(1,2), 3))
v5 <- c(rep(c(1,2), 3, each=2))
cat(v1, v2, v3, v4, v5)
```

(indexing)

```{r}
v1[1]
v1[-2]
v1[v1%%2 == 1]
v1[2:4]
v1[c(1, 5)]
```

(vector functions)

```{r}
v1 <- c(-1, 0, 2, 4, 15, 62, -7)
v2 <- c(-2, 8, 5, 3, 1, -1, 9)

mean(v1)
sum(v1)
median(v1)
range(v1)
sd(v1)
var(v1)
cov(v1, v2)
cor(v1, v2)
min(v1)
max(v1)
```


(vectorized operations)

```{r}
v1 <- c(-1, 0, 2, 4, 15, 62, -7)
v2 <- c(-2, 8, 5, 3, 1, -1, 9)
v3 <- c(1, 56, 15, 62, 143, 23, 18)
v4 <- c(-1.5, -7.3, -7.7, 1.5, 7.3, 7.7)

sqrt(v3)
log(v3, base=10)
abs(v2)
round(v4)
trunc(v4)
```



### list

(key, value) pairs로 데이터 저장.

```{r}
lkst = list(vec=v1, vl=a2)
lkst

lkst$vec
lkst$vec[1]

lkst$vl
```


### matrix

matrix : 2-dim vector.

```{r}
dt <- c(11, 12, 13, 14, 15, 21, 22, 23, 24, 25, 31, 32, 33, 34, 35)

mtx1 <- matrix(dt, nrow=5, ncol=3)
mtx1
mtx2 <- matrix(dt, nrow=3, ncol=5)
mtx2
mtx3 <- matrix(dt, nrow=5, ncol=3, byrow=TRUE)
mtx3
mtx4 <- matrix(dt, nrow=3, ncol=5, byrow=TRUE)
mtx4
mtx5 <- matrix(dt, nrow=5, ncol=3, byrow=TRUE, dimnames=list(c('i1', 'i2', 'i3', 'i4', 'i5'), c('c1', 'c2', 'c3')))
mtx5
mtx6 <- dt
dim(mtx6) <- c(5, 3)
mtx6
```

(matrix operations)

```{r}
dim(mtx5)
nrow(mtx5)
ncol(mtx5)

mtx5[4, 2]
mtx5[4,]
mtx5[,2]

rownames(mtx5)
colnames(mtx5)

rowSums(mtx5)
colSums(mtx5)

rowMeans(mtx5)
colMeans(mtx5)
```


### data.frame

```{r}
df1 <- data.frame(v1, v2, v3, f1=c('F', 'F', 'F', 'M', 'U', 'M', 'U'))
df1
df2 <- data.frame(v1, v2, v3, f1=c('F', 'F', 'F', 'M', 'U', 'M', 'U'), 
                  stringsAsFactors=TRUE)
df2
```


### factor

```{r}
data_vector <- c('M', 'F', 'U', 'F', 'F', 'U', 'M', 'F')
levels_vector <- c('M', 'F', 'U')
display_vector <- c('Male', 'Female', 'Unknown')

factor(data_vector, levels_vector, display_vector)
```

```{r}
data_vector <- c('M', 'F', 'U', 'F', 'F', 'U', 'M', 'F')
levels_vector <- c('M', 'F', 'U')
display_vector <- c('Good', 'Soso', 'Bad')

factor(data_vector, levels_vector, display_vector, ordered = FALSE)
```


---

## Flow Control : Conditionals

### if - else if - else 

```{r}
score <- 20
if (score > 90) {
  print('A')
} else if (score > 80){
  print('B')
} else {
  print('F')
}
```


### ifelse

```{r}
v1 <- c(1, 0, 1, 2, 6, 3, 2, 1, 5, 6)
ifelse(v1 > 5, TRUE, FALSE)
```


### switch

```{r}
course <- '1'
switch(course, '0' = 'A', '1' = 'F')
```

vectorized 연산 불가. 





## Flow control : Iteration

### for

```{r}
for (i in 1:10) {
  print(i)
}
```

### while

```{r}
i <- 1
while (i <= 10) {
  print(i)
  i <- i + 1
}
```

```{r}
i <- 1
while (i <= 10) {
  print(i)
  i <- i + 1
}
```


### repeat

```{r}
i <- 1
repeat{
  print(i)
  i <- i + 1
  if (i > 10) {
    break
  }
  }
```


### loop control : 


(break)

```{r}
for (i in 1:10) {
  print(i)
  if (i%/%2 == 0) {
    break
  }
}
```


(next)

```{r}
for (i in 1:10) {
  if (i == 5) {
    next
  }
    print(i)
}
```



---

## User Function

```{r}
userfunc <- function(arg1, arg2){
  temp <- arg1 - arg2
  return((arg1+arg2)*temp)
}

userfunc(4, 7)
```

















---

# File I/O

(Working Directory)

```{r}
getwd()
#setwd('C:/Users/Na/Documents/GitHub/R-Practice/R_Basics')
```


(absolute path)

```{r}
df <- read.csv("C:/Users/Na/Documents/GitHub/R-Practice/R_Basics/R_Basics_csvread.csv", header = F)
df
comment = '
Generally, it is not recommended to use absolute path
because of reproductivity issue.
'
```


(relative path)

```{r}
df <- read.csv("R_Basics_csvread.csv", header = F)
df
```










---

# R Base - Visualization

## visualization - univariate

```{r}
gen1 <- factor(c('F','F','M','M','F'))
gen2 <- c('F','F','M','M','F')
score <- c(44, 55, 66, 77, 55)
t1 <- cbind(gen1, score)
t2 <- cbind(gen2, score)
t1
t2
```

```{r}
hist(score)
hist(score, breaks = 10)
hist(score, freq = F)#비율
hist(score, breaks = c(5,15,25,35,45,55,65,75,85))
hist(score, breaks = seq(5, 105, by = 2))
plot(score, type = 'h')
```

```{r}
stem(score)
boxplot(score)
boxplot(score, horizontal = T)
boxplot(score~gen1)#bivarate
```

```{r}
table(gen1)
table(score)
fivenum(score)
summary(score)
range(score)
IQR(score)
mean(score)
var(score)
sd(score)
median(score)
sum(score)
min(score)
max(score)
round(sd(score))
round(sd(score), digits = 2)
rank(score)
```


## Visualization - bivariate

```{r}
plot(score, gen1, main = 'nyan', sub = 'nnyan')
plot(score, gen1, main = 'nyan', sub = 'nnyan', xlab = 'kodon', ylab = 'GenG', type = 'l')
plot(score, gen1, main = 'nyan', sub = 'nnyan', xlab = 'kodon', ylab = 'GenG', type = 'c')
plot(score, gen1, main = 'nyan', sub = 'nnyan', xlab = 'kodon', ylab = 'GenG', type = 'n')
plot(score, gen1, main = 'nyan', sub = 'nnyan', xlab = 'kodon', ylab = 'GenG', type = 'b')
plot(score, gen1, main = 'nyan', sub = 'nnyan', xlab = 'kodon', ylab = 'GenG', cex = 0.5)
plot(score, gen1, main = 'nyan', sub = 'nnyan', xlab = 'kodon', ylab = 'GenG', cex = 2)
plot(score, gen1, main = 'nyan', sub = 'nnyan', xlab = 'kodon', ylab = 'GenG', cex = 2, pch = 16)
```



## multi graph

```{r}
par(mfrow = c(2,3))
for (i in 1:6) {
  a <- seq(-3, 3, by = 0.1)
  b <- dnorm(a)
  c <- seq(-3, 3, by = 1)
  d <- dnorm(c)
  plot(a, b, pch = 16, cex = 0.2)
  lines(c, d)
}
dev.off
```



## Q-Q plot

```{r}
dat <- rt(20, df=10)
qqnorm(dat, pch = 16, cex = 0.2) ; qqline(dat)
comment = 'lines와 마찬가지로 이미 존재하는 그래프 위에 그리는 것.'
```





---

# Rbase : useful statistics function

## `sample()`

```{r}
sample(1:100, 20, replace = T)
sample(1:100, 20, replace = F)
#sample(x, n, replace = F/T) : x에서 size n vector를 추출.
```





## `factorial()`, `choose()`

```{r}
choose(12,3)
factorial(12)
```






## distribution call

```{r}

```







### normal distribution

```{r}
dnorm(0, mean = 0, sd = 1)#point 0에서의 함숫값
pnorm(0, mean = 0, sd = 1)#point 0에서의 분포함숫값
qnorm(0.5, mean = 0, sd = 1)#z_0.5가 되는 x값
pnorm(1, lower.tail = F)#클 확률 구함
qnorm(0.1587, lower.tail = F)#큰 값이 0.3413인 point
```

```{r}
a <- seq(-3, 3, by = 0.1)
b <- dnorm(a)
plot(a, b, type = 'l', xlab = 'x', ylab = 'y', main = 'probablilty density function of standard normal distribution')
```



### Binomial distribution

```{r}
dbinom(5, 10, 0.5)
pbinom(5, 10, 0.5)
qbinom(0.5, 10, 0.5)
```



























