---
title: "DA Lab7_HW"
author: "Na SeungChan"
date: "`r Sys.Date()`"
mainfont : NanumGothic
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(latex2exp)
library(brms)
library(broom)
library(MASS)
library(boot)
```


해당 강의 실습문항을 꽤나 못 풀었으나 일단 제출합니다. 일단 풀어 보고 조만간 메일 등으로 질문드리겠습니다.


## Exercises

### Exercise 1

(a) Using the `affair` data set, select three variables which you think are related to the probability of cheating. Perform a proper regression model and explain the result.

```{r}
affairs_df <- read_csv('affairs.csv') %>%
  mutate(cheater = ifelse(affairs > 0, TRUE, FALSE))

affairs_glm_bin <- glm(cheater ~ children + education + rating, family = binomial(link = "logit"), data = affairs_df)

summary(affairs_glm_bin)
```



(b) Compare the deviance with the model we fitted in class.

Residual deviance와 AIC 모두 감소하였다.



(c) Consider that there is a new individual with the three variables you chose above. Randomly select each variable and guess his or her probability of cheating.

```{r, eval=FALSE}

predict.glm(affairs_glm_bin, newdata = data_frame(sample(c(TRUE, FALSE), 1, 1), sample(seq(9, 20, 1), 1, 1), sample(seq(1, 5, 1), 1, 1)))

```




### Exercise 2

(a) Using the `low birth weight infants` data set, select three variables which you think are related to the number of visits to the doctor. Perform a proper regression model and explain the result. 

```{r}
lbw_df <- read_csv('lbw.csv')
lbw_df

lbw_glm <- glm(ftv ~ age + smoke + lwt, family = poisson(link = 'log'), data = lbw_df)

summary(lbw_glm)
```



(b) Compare the deviance with the model we fitted in class.

252.96이었던 residual deviance가 250.25로 감소하였으나, AIC는 468.33으로 증가하였다.



(c) Consider that there is a new individual with the three variables you chose above. Randomly select each variable and guess her number of visits to the doctor.



### Exercise 3
Perform a forward selection estimating the probability of cheating with your own criterion. State the best model and explain the result.


### Exercise 4
Build your model that estimates the number of visits to the doctor. State the model and explain the result.



### Exercise 5

(a) Using the `affair` data set, fit the following five models and compare the result.
- Intercept model
- Model using 'age', 'gender', 'education', 'yearmarried', 'occupation' variables only
- Model using 'age', 'gender', 'education', 'yearmarried', 'occupation' variables and their two way interactions
- Model using 'age', 'gender', 'education', 'yearmarried', 'occupation' variables and their two and three way interactions.
- Full model using 'age', 'gender', 'education', 'yearmarried', 'occupation' variables and their all interactions.


```{r}
affairs_glm_bin1 <- glm(cheater ~ 1, family = binomial(link = "logit"), data = affairs_df)
summary(affairs_glm_bin1)

affairs_glm_bin2 <- glm(cheater ~ (age + gender + education + yearsmarried + occupation)^2, family = binomial(link = "logit"), data = affairs_df)
summary(affairs_glm_bin2)

affairs_glm_bin3 <- glm(cheater ~ (age + gender + education + yearsmarried + occupation)^3, family = binomial(link = "logit"), data = affairs_df)
summary(affairs_glm_bin3)

affairs_glm_bin4 <- glm(cheater ~ age*gender*education*yearsmarried*occupation, family = binomial(link = "logit"), data = affairs_df)
summary(affairs_glm_bin4)
```



(b) Using the LOOCV, explain which model shows the best performance.

```{r}
cver_bin1 <- cv.glm(affairs_df, affairs_glm_bin1)
cver_bin1$delta

cver_bin2 <- cv.glm(affairs_df, affairs_glm_bin2)
cver_bin2$delta

cver_bin3 <- cv.glm(affairs_df, affairs_glm_bin3)
cver_bin3$delta

cver_bin4 <- cv.glm(affairs_df, affairs_glm_bin4)
cver_bin4$delta
```

시행 매우 느려... 결과가 뭔가 이상하다.



### Exercise 6

(a) Using the `low birth weight infants` data set, fit the following five models and compare the result.
- Intercept model
- Model using 'age' and 'lwt' variables only
- Model using 'age' and 'lwt' variables and their quadratic terms
- Model using 'age' and 'lwt' variables and their quadratic/cubic terms
- Model using 'age' and 'lwt' variables and their interaction.

```{r}
lbw_m1 <- glm(ftv ~ 1, family = poisson(link = 'log'), data = lbw_df)
summary(lbw_m1)

lbw_m2 <- glm(ftv ~ poly(age, lwt, degree = 2), family = poisson(link = 'log'), data = lbw_df)
summary(lbw_m2)

lbw_m3 <- glm(ftv ~ poly(age, lwt, degree = 3), family = poisson(link = 'log'), data = lbw_df)
summary(lbw_m3)

lbw_m4 <- glm(ftv ~ age*lwt, family = poisson(link = 'log'), data = lbw_df)
summary(lbw_m4)
```



(b) Using the K-fold cross-validation, explain which model shows the best performance.

```{r}

```

